---
title: "Math_Final_Project"
author: "Mauricio Sanchez"
date: "November 30, 2015"
output: html_document
---
#Math –Final Project
Objective: To use the hflights package or another large (greater than 100,000 observations and 20 variables) dataset
of your choosing, pick two quantitative variables (e.g., Arrival Delay and Departure Delay). 

#1
Invoke the hflights package and display the head data of the entire data set.
```{r}
require (hflights);
require(ggplot2);
warning(FALSE);
```

#2
Define one of these variables as the random variable X and the other as the random variable Y. One of these variables
must be skewed to the right (positively)
```{r, echo=TRUE}
x <- hflights$ArrDelay;
y <- hflights$DepDelay;

summary(x);
summary(y);

# NAs = 3622 
condTotal <- length(x) - 3622;
condTotal;
```

#3
I have chosen the Arrival Delay as the variable that is skewed to the right positively with mean greater than the median.
```{r, echo=TRUE}
ggplot(data = hflights) + geom_density(aes(x=hflights$ArrDelay), fill="blue") + ggtitle("Arrival Delay Density");
```

#4 Probability 
Calculate as a minimum the below probabilities a through d. Assume the small letter "x" is
estimated as the 3d quartile of the X variable, and the small letter "y" is estimated as the 2d quartile of
the Y variable. Interpret the meaning of all probabilities. In addition, make a table of counts as shown
below. a. P(X>x | Y>y) b. P(X>x, Y>y) c. P(X<x | Y>y) d. P(X<x | Y>y).

x is estimated as the third quartile of the X variable -> 9
y is estimated as the second quartile of Y variable -> 0

```{r, echo=TRUE}
#a P(X>x | Y>y)
val1 <- nrow(subset(hflights, x > 9 & y > 0));
val1;

#condTotal = 223874
(val1/condTotal)/0.5;
0.25*0.50;

#b. 
val2 <- nrow(subset(hflights, x <= 9 & y > 0));
val2;
(val2/condTotal)/0.5;

#c 
val3 <- nrow(subset(hflights, x <= 9 & y <= 0));
val3;
(val3/condTotal)/0.5;

#d
val4 <- nrow(subset(hflights, x > 9 & y <= 0));
val4;
(val4/condTotal)/0.5;

#table
total1 <- sum(val2,val3);
total2 <- sum(val4,val1);

a <- c('<=3d quartile', '>3d quartile','Total');
b <- c(val2,val3,(total1));
c <- c(val4,val1,(total2));
d <- c((val2+val4),(val3+val1),(total1+total2));
levels <- factor(c("x/y","<=2d quartile",">2d quartile","Total"));
resultsTbl <- data.frame('x/y'=a,
                         '<=2d quartile'=b,
                         '>2d quartile'=c,
                         'Total'=d);
resultsTbl;
```

Does P(A|B)=P(A)P(B)? 
```{r, echo=TRUE}
A = total2;
B = total1;

PA = A/condTotal;
PB = B/condTotal;

PA_B = PA/PB;
PA_B == PA*PB;
```

Evaluate by running a Chi Square test for association on x and y. 
```{r, echo=TRUE}
chisq.test(x,y);
```

#5 Descriptive and Inferential Statistics
Provide a scatterplot of the two variables x and y
```{r, echo=TRUE}
plot(x, y, xlab="Arrival Delays", ylab="Departure Delays", main="Arrival & Departure Delays", pch=2, cex.main=1.5, frame.plot=FALSE);
```

Provide a 95% CI for the difference in the mean of the variables x and y
```{r, echo=TRUE}
t.test(y,x);
```

Derive a correlation matrix for two of the quantitative variables you selected. Test the
hypothesis that the correlation between these variables is 0 and provide a 99% confidence interval. Discuss the meaning of your analysis.
```{r, echo=TRUE}
data <- data.frame(x,y);
matrix <- cor(data);
cor.test(y,x,conf.level=0.99);
```
This above shows 95% confidence that the difference in the means of x and y will be between 2.176342 and 2.524892. I conclude that correlation may exist based on the 99 percent confidence value. True correlation is not equal to 0. 

