---
title: "R_Final_Project"
author: "Mauricio Sanchez"
date: "November 29, 2015"
output: html_document
---
#R Bridge –Final Project
Objective: To analyze a subset of the hflights data to draw a parallel between airlines that have the least and the most delays. This will help one make an intelligent decision when choosing airlines flying out of Houston airports.

#1
Invoke the hflights package and display the head data of the entire data set.
```{r}
require (hflights);
require(ggplot2);
require(sqldf);
warning(FALSE);
```

#2
Format Date
```{r, echo=TRUE}
formatdate <- paste(hflights$Month, hflights$DayofMonth, hflights$Year, sep="-");
hflights$Date <- formatdate;
hflights$Date <- as.Date(formatdate, "%m-%d-%Y");
```

#3
Translate Week Days
```{r, echo=TRUE}
hflights$DayOfWeek <- sub(1, 'Monday', hflights$DayOfWeek)
hflights$DayOfWeek <- sub(2, 'Tuesday', hflights$DayOfWeek)
hflights$DayOfWeek <- sub(3, 'Wednesday', hflights$DayOfWeek)
hflights$DayOfWeek <- sub(4, 'Thursday', hflights$DayOfWeek)
hflights$DayOfWeek <- sub(5, 'Friday', hflights$DayOfWeek)
hflights$DayOfWeek <- sub(6, 'Saturday', hflights$DayOfWeek)
hflights$DayOfWeek <- sub(7, 'Sunday', hflights$DayOfWeek)
```

#3
Quick Snapshot of the Data
```{r, echo=TRUE}
head(hflights);
```

#4
Choose a subset of the data.
```{r, echo=TRUE}
dd1 <- sqldf("SELECT UniqueCarrier, COUNT(ArrDelay) AS ArrivalDelays FROM hflights WHERE ArrDelay > 0 GROUP BY UniqueCarrier ORDER BY 2 DESC");
dd2 <- sqldf("SELECT UniqueCarrier, COUNT(Distance) AS Flights FROM hflights WHERE Distance > 0 GROUP BY UniqueCarrier ORDER BY UniqueCarrier ASC");
dd3 <- merge(dd1, dd2, by="UniqueCarrier");
dd3["Ratio"] <- round(((dd3$ArrivalDelays/dd3$Flights)*100), digits=0);
```

#5
Show Arrival Delays by Airline. 
```{r, echo=TRUE}
print(dd3);
```

#6
The airline with the highest ratio of arrival delays is Frontier Flight (F9) with a 55 percent delay ratio. The airline with the lowest arrival delays is American Airlines (AA) with 30 percent.
```{r, echo=TRUE}
ggplot(hflights, aes(x = UniqueCarrier, y = ArrDelay)) + geom_boxplot() + ggtitle("Arrival Delays by Airline by Number of Flights");
ggplot(data = dd3) + geom_density(aes(x=ArrivalDelays), fill="grey50") + ggtitle("Arrival Delay Density");
ggplot(data=dd3) + geom_bar(aes(x=Ratio), bin=5) + ggtitle("Arrival Delay Ratio");
```

#7
Let's look at the number of flights vs. delay ratio. This graph clearly shows the airlines with the highest and lowest ratio of arrival delays. 
```{r, echo=TRUE, warning=TRUE}
qplot(data=dd3, Flights, Ratio, color = UniqueCarrier) + facet_grid(~UniqueCarrier) + ggtitle("Arrival Delay Ratio by Airline");
```

#8 
Let's analyze arrival delays by day of the week. This graph suggest that there is a correlation between departure and arrival delays increasing the delays as the number of flights increase. This could be caused as a result that the days when there are more fligths there are more delays in departures and arrivals. 
```{r, echo=TRUE, warning=FALSE}
qplot(data=hflights, ArrDelay, DepDelay, color = UniqueCarrier) + facet_grid(~DayOfWeek) + ggtitle("Arrival vs. Departure Delays by WeekDay");
```

#9 Questions
1. What are the days with the least amount of departure and arrival delays? Tuesdays and Thursdays are the days with the least amount of delays.

2. What is the best airline to fly with the least amount of arrival delays? I would choose American Airlines to fly into Houston.

3. What days would you avoid flying in or out of the houston airports? I would totally avoid the weekends to fly in or out of Houston. The number of flights increase on weekends and therefore the number of delays.